---
title: "Assignment 2 - The tidyverse"
author: "Sangeeth Selvaraju"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---
  
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```


***

```{r, include = T}
# LOAD THE PACKAGES YOU ARE USING IN THIS CODE CHUNK library(nameofpackage)

library(legislatoR)
library(tidyverse)
library(ggplot2)
library(tibble)
library(lubridate)
library(table1)
?legislatoR()
```

<br>

### Getting started with the Comparative Legislators Database

The Comparative Legislators Database (CLD) includes political, sociodemographic, career, online presence, public attention, and visual information for over 45,000 contemporary and historical politicians from ten countries (see the [corresponding article](https://www.cambridge.org/core/journals/british-journal-of-political-science/article/comparative-legislators-database/D28BB58A8B2C08C8593DB741F42C18B2) for more information). It can be accessed via `legislatoR` - an R package that is available on both [CRAN](https://cran.r-project.org/web/packages/legislatoR/index.html) and [GitHub](https://github.com/saschagobel/legislatoR).

Before you start with the following tasks, skim the tutorial to make yourself familiar with the database. You find it [here](https://cran.r-project.org/web/packages/legislatoR/vignettes/legislatoR.html).

For the following tasks, you will work with ONE of the legislatures. The legislature you'll work with depends on your prename:

| Your prename starts with... | Legislature | Code |
|---------|-------|-------|
| A-C | Austrian Nationalrat | `aut` |
| D-F | Canadian House of Commons | `can` |
| G-I | Czech Poslanecka Snemovna | `cze` |
| J-L | Spanish Congreso de los Diputados | `esp` |
| M-O | French Assemblée | `fra` |
| P-R | German Bundestag | `deu` |
| S-U | Irish Dail | `irl` |
| V-X | UK House of Commons | `gbr` |
| Y-Z | US House of Representatives  | `usa_house` |

The following tasks will be based on data from the `core` and `political` tables of the database.

<br>

***

### Task 1 - Fun with descriptives [8 points in total]

a) What's the overall share of female legislators in the entire dataset? [1 point]
Answer) Ireland has 9% female legislators in the dataset
```{r}
# using filter from tidyverse to get this
##getting just Irish Dail core sociodemographic data

irl_house_core <- get_core(legislature = "irl")

irl_house_core_1 <- irl_house_core %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(prop =n /sum(n, na.rm = T))

```

b) How many (both in absolute and relative terms) legislators died in the same place they were born in? [1 point]
Answer) 42 legislators died and were born in the same place
```{r}
# 
irl_house_core_2 <- irl_house_core %>% 
  mutate(birth_death = ifelse(birthplace == deathplace, yes = TRUE, no = FALSE)) %>% 
  group_by(birth_death) %>%
  summarise(n = n()) %>%
  mutate(proportion = n/sum(n))

# add column to return 'TRUE' if birth and death place coordinates are the same
```

c) Create a new variable `age_at_death` that reports the age at which legislators died. Then, plot the distribution of that variable for the entire dataset. [2 points]

```{r}
class(irl_house_core$death)
## using package lubridate to change the birth and death format of variables to date
## creating new variable and calculating age at death in years

irl_house_core <- irl_house_core %>% mutate(death_updated = as_date(death), 
                              birth_updated = as_date(birth), 
                              age_at_death = as.numeric((death_updated -          birth_updated)/365.25))

##ggplot histrogram to show the age at death of legislators
ggplot(irl_house_core, aes(x= age_at_death)) +
   geom_histogram(bandwith = 2, color = "black" ,  fill = "grey") +
  labs(x = "Legislators age at death(Years)", y = "Number of legislators")+
  theme_economist() + scale_colour_economist()
```

d) Provide one single table that provides the 5 legislators with the highest and lowest age at death, ignoring ties. [2 points]

```{r}
top_5 <- irl_house_core%>% 
  select(name, age_at_death) %>%
  slice_max(n = 5, order_by = age_at_death) 

bottom_5 <- irl_house_core %>% 
  select(name, age_at_death) %>%
  slice_min(n = 5, order_by = age_at_death) 

# Bind together, HighandLow_Age_at_death_Legislator shows a table of 5 legislators with the highest and lowest age at death
HighandLow_Age_at_death_Legislator  <- as_tibble((bind_rows(top_5, bottom_5)))
```

e) Find out who has got the longest name and who the shortest (both in terms of number of characters, ignoring whitespace)! [2 points]

```{r}
##using package stringr
irl_house_core <- irl_house_core %>% 
   mutate(name_lenght = as.character(name)) %>% 
   
       

```

<br>

***

### Task 2 - Exploring trends in the data [8 points in total]

a) Using data from all sessions, provide a plot that shows the share of female legislators by session! [3 points]

```{r}
##loaded the politcal dataset so at to get session information
irl_house_pol <- get_political(legislature = "irl")

## left joined the core dataset to the political dataset because politcal dataset has ~5000 something observation and core only ~1400 so i wanted to retain all the data, the other way around would have lost data from politcal dataset
irldf <- left_join(irl_house_pol,irl_house_core)
```

```{r}
ggplot(irldf) +
   geom_bar(position = "stack", aes(x=session, y=sex),
           bandwith = 2, color = "black", fill = "pink")+
           labs(x = "Legislative sessions" , y = "number of female legislators") +
          theme_economist() +
         scale_colour_economist()
```

b) Explore another relation of variables in the dataset. Your analysis should (i) use data from both the `core` and the `political` table and (ii) feature data from several sessions. Also, offer a brief description and interpretation of your findings! [5 points]

```{r}
# YOUR CODE HERE  ##regression
```

<div class = "answer">
YOUR ANSWER SHOULD GO HERE
</div>

<br>

***

### Task 3 - Fun with functions [3 points in total]

a) Program a function ultimate_answer() that always returns the number 42, and prove that it works! [1 point]

```{r}
# YOUR CODE HERE
```

b) Program a function normalize() that normalizes a numeric vector x to mean(x) = 0 and sd(x) = 1! Then, apply it to all numeric variables in the mtcars dataset! [2 points]

```{r}
# YOUR CODE HERE
```



